---
title: "A3 God Dataset"
author: "Swetha KV"
date: "2025-11-01"
categories: [God]
format: html
---

```{r}
library(tidyverse)
library(mosaic)
library(vcd)
library(visStatistics)
library(ggformula)
library(infer)
library(openintro)
```

```{r}
god <- read.csv("kadavulirukanda.csv")
glimpse(god)
```

```{r}
god <- god%>% 
 tidyr::drop_na()
```

```{r}
visdat::vis_dat(god)
```

```{r}
god %>%
  group_by(Believe_in_God) %>%
  count()
```

```{r}
god %>%
  group_by(Gender) %>%
  count()
```

```{r}
god %>%
  summarize(prop = mean(Believe_in_God == "Yes"), n = n())
```

```{r}
god_table <- vcd::structable(Gender ~ Believe_in_God, data = god)
god_table
```

```{r}
god %>%
  gf_bar(~Believe_in_God, fill = ~Gender) %>%
  gf_labs(
    x = "Believe in God",
    title = "Proportion of Students saying they believe in god"
  )
```
### Inference 

- The bar graph shows a moderate distribution of belief, indicating coexistence of both religious and secular perspectives among students.

```{r}
mosaic::binom.test(~Believe_in_God, data = god, success = "Yes")
```

```{r}
god %>%
  drop_na() %>%
  specify(response = Believe_in_God, success = "Yes") %>%
  generate(reps = 999, type = "bootstrap") %>%
  calculate(stat = "prop") %>%
  get_ci(level = 0.95)
```

# Conlcusion

### Hypothesis

Numerical hypothesis: The proportion of students that believe in God is equal to 0.5 (no difference and equally to believe and disbelieve).

Alternative hypothesis: The proportion of students that believe in God is not equal to 0.5 (there is a difference between believe and disbelieve).

### Inference

- Since the p-value (0.1173) > 0.05, we fail to reject the null hypothesis.

- There isn’t strong statistical evidence that the proportion of students who believe in God differs significantly from 50%.

- The sample proportion (63%) does lean higher than 0.5, which means tendency toward belief that become significant.

- The 95% confidence interval (0.469 to 0.779) means we are 95% confident that the true population proportion of students who believe in God lies between 46.9% and 77.9%.

- Since 0.5 (the neutral value) lies within this interval, we can’t claim a significant deviation from 50%.

- The unequal gender sample (23 female, 18 male) have slightly biased the overall proportion upward because the graph is showing that females tend to express higher belief rates. A gender-balanced sample might give us a slightly lower or more neutral proportion.

