---
title: "A3 God Dataset"
author: "Swetha KV"
date: "2025-11-01"
categories: [A3]
format: html
---

# Do people believe in God?

## Setting up R Packages

```{r}
library(tidyverse)
library(mosaic)
library(vcd)
library(visStatistics)
library(resampledata)
library(ggformula)
library(infer)
library(openintro)
```

## Read Data

```{r}
god <- read.csv("kadavulirukanda.csv")
glimpse(god)
```

## Cleaning Data

```{r}
god <- god%>% 
 tidyr::drop_na()
```

```{r}
visdat::vis_dat(god)
```

## Summerise: Examine Data

```{r}
god %>%
  group_by(Believe_in_God) %>%
  count()
```

```{r}
god %>%
  group_by(Gender) %>%
  count()
```

```{r}
god %>%
  summarize(prop = mean(Believe_in_God == "Yes"), n = n())
```

```{r}
god_table <- vcd::structable(Gender ~ Believe_in_God, data = god)
god_table
```

## Visualising Data

```{r}
god %>%
  gf_bar(~Believe_in_God, fill = ~Gender) %>%
  gf_labs(
    x = "Believe in God",
    title = "Proportion of Students saying they believe in god"
  )
```

### Inference

-   The bar graph shows a moderate distribution of belief, indicating coexistence of both religious and secular perspectives among students.

```{r}
god %>% 
gf_bar(~Gender, fill = ~Believe_in_God,
       position = "dodge") %>% 
gf_labs(x="gender",
        y="count",
        title=" Genders Vs Beliefs in God")
```

```{r}
god %>%
  gf_bar(~Believe_in_God, fill = ~Gender, position = "fill") %>%
  gf_labs(
    x = "Believe in God",
    title = "Gender difference believe in god"
  )
```

## Propotion Test

```{r}
mosaic::binom.test(~Believe_in_God, data = god, success = "Yes")
```

### Inference

-   Since the p-value (0.1173) \> 0.05, we fail to reject the null hypothesis.

-   The sample proportion (63%) does lean higher than 0.5, which means tendency toward belief that become significant.

### Hypothesis

Numerical hypothesis: The proportion of students that believe in God is equal to 0.5 (no difference and equally to believe and disbelieve).

Alternative hypothesis: The proportion of students that believe in God is not equal to 0.5 (there is a difference between believe and disbelieve).

```{r}
god %>%
  drop_na() %>%
  specify(response = Believe_in_God, success = "Yes") %>%
  generate(reps = 999, type = "bootstrap") %>%
  calculate(stat = "prop") %>%
  get_ci(level = 0.95)
```

### Inference

-   The 95% confidence interval (0.469 to 0.779) means we are 95% confident that the true population proportion of students who believe in God lies between 46.9% and 77.9%.

```{r}
contingency_table <- vcd::structable(Believe_in_God~ Gender,
  data = god
)
contingency_table
```

```{r}
vcd::mosaic(Gender ~ Believe_in_God, data = god, 
            direction = "v", gp = shading_max, legend = TRUE)

vcd::mosaic(Gender ~ Believe_in_God, data = god, 
            type = "expected", direction = "v", gp = shading_max, 
            legend = TRUE)

vcd::assoc(Gender ~ Believe_in_God, 
           data = god, 
           gp = shading_max, 
           legend = TRUE) 
```

```{r}
xq_test_object <- chisq.test(contingency_table)
xq_test_object
```

### Inference

Null Hypothesis: There is no association between gender and in belief in God Alternative Hypothesis: Null Hypothesis is true

## Conclusion

-   The unequal gender sample (23 female, 18 male) have slightly biased the overall proportion upward because the graph is showing that females tend to express higher belief rates. A gender-balanced sample might give us a slightly lower or more neutral proportion.
