---
title: "A3 TV Shows Dataset"
author: "Swetha KV"
date: "2025-10-31"
categories: [A3]
format: html
---
# Friends vs Modern Family vs Big Bang Theory!!!

## Setting up R Packages

```{r}
library(tidyverse)
library(ggformula)
library(mosaic)
library(broom)
library(infer)
library(supernova)
library(ggstatsplot)
library(ggcompare)
library(patchwork)
library(ggprism)
library(paletteer)
library(DescTools)
library(latex2exp)

```

## Read Data
```{r}
ratings <- read.csv("tvseries.csv")
glimpse(ratings)
```
```{r}
ratings <- ratings %>% drop_na()
```

```{r}
visdat::vis_dat(ratings)
```

## Data Munging

### Convert from wide to long format

```{r}
names(ratings)
```


```{r}
ratings_long <- ratings %>%
  pivot_longer(
    cols = c(Friends, Modern.Family, Big.Bang.Theory),
    names_to = "Show",
    values_to = "Rating"
  ) %>%
  drop_na() %>%
  mutate(
    Show = factor(
      Show,
      levels = c("Friends", "Modern.Family", "Big.Bang.Theory"),
      labels = c("Friends", "Modern Family", "Big Bang Theory")
    ),
    Gender = factor(Gender, levels = c("Male", "Female"))
  )

ratings_long

```

## Visualising Data

```{r}
gf_jitter(Rating ~ Show, color = ~Show, width = 0.2, data = ratings_long, size = 3, alpha = 0.5) %>%
  gf_labs(title = "Opinion Scores by Show", x = "Show", y = "Rating") %>%
  gf_hline(yintercept = ~ mean(Rating), color = "grey") %>%
  gf_annotate(geom = "text", label = "Overall Mean", x = 3, y = mean(ratings_long$Rating) + 0.2, size = 4)

```


```{r}
ratings_long %>%
  gf_boxplot(Rating ~ Show, orientation = "x", fill = ~ Show) %>%
  gf_labs(
    title = "Rating vs Show",
    x = "TV Shows",
    y = "Ratings"
  ) %>%
  gf_hline(yintercept = ~ mean(Rating), color = "red") %>%
  gf_annotate(
    geom = "text",
    label = "Overall Mean",
    x = 3,
    y = mean(ratings_long$Rating) + 0.3
  ) %>% 
  gf_theme(theme_minimal()) %>% 
  gf_refine(scale_fill_brewer(palette = "Set2"))

```
### Inference

- Modern Family received the highest ratings, with most scores around 8–9.
- Friends ratings are moderate, clustering near 7.
- Big Bang Theory got the lowest ratings, mostly between 6–7.
- The overall mean (blue line) lies near 7.2, showing that the three shows differ clearly in how people rated them.

```{r}
gf_point(Rating ~ Show, color = ~Gender, data = ratings_long, 
         position = position_dodge(width = 0.5)) %>% 
  gf_line(Rating ~ Show, color = ~Gender, data = ratings_long, 
          stat = "summary", fun.y = "mean", 
          position = position_dodge(width = 0.5))
```
### Inference

- Both males and females gave a wide range of ratings for all three shows.
- However, females tended to give slightly higher ratings overall, especially for Modern Family.
- There is no extreme difference between genders, but males showed more variation in their ratings.


## ANOVA Test

```{r}
ratings_anova <- aov(Rating ~ Show, data = ratings_long)
summary(ratings_anova)

supernova::supernova(ratings_anova)
```

```{r}
ratings_anova <- aov(Rating ~ Show, data = ratings_long)
ratings_anova
```
### Inference

- Null Hypothesis Rejected: We confidently reject the idea that all three shows have the same true average rating.

- Alternative Hypothesis: This means we accept the alternative hypothesis, concluding that there is a significant difference in the mean ratings among the TV shows.

```{r}
ratings_anova_1 <- aov(Rating ~ Show + Gender, data = ratings_long)
summary(ratings_anova)

supernova::supernova(ratings_anova_1)
```
### Inference
- Mean ratings of the shows are significantly different AND ratings given by different genders are also significantly different.

```{r}
ratings_supernova <-
  supernova::pairwise(ratings_anova_1,
    correction = "Bonferroni", # Try "Tukey"
    alpha = 0.05, # 95% CI calculation
    var_equal = TRUE, # We'll see
    plot = T
  )
```
### Inferemce

- Modern Family received ratings that are significantly higher than the ratings for Friends.

- Big Bang Theory and Friends ratings are not significantly different from each other.

- Modern Family received ratings that are significantly higher than the ratings for Big Bang Theory.

- The difference between females and males is not statistically significant.

- This means we cannot clearly say one gender is different from the other based on this test.

```{r}
ratings_supernova$Show
```

### Inference

- Modern Family received ratings that are significantly higher than the ratings for Friends.

- Big Bang Theory and Friends ratings are not significantly different from each other.

- Modern Family received ratings that are significantly higher than the ratings for Big Bang Theory.


```{r}
supernova::supernova(ratings_anova_1)
```

```{r}
supernova::equation(ratings_anova)
```

```{r}
ratings_anova <- aov(Rating ~ Show, data = ratings_long)

ratings_long %>%
  mutate(fitted = fitted(ratings_anova)) %>%
  gf_jitter(
    Rating ~ Show,
    width = 0.2,
    alpha = 0.6,
    color = ~Show,
    data = .
  ) %>%
  gf_summary(
    group = ~1,
    fun = "mean",
    geom = "line",
    colour = "lightblue",
    lty = 1,
    linewidth = 2
  ) %>%
  gf_point(
    fitted ~ Show,
    color = ~Show,
    shape = 2,
    size = 6
  ) %>%
  gf_segment(
    fitted + fitted ~ 0 + Show,
    linetype = 2,
    color = ~Show
  ) %>%
  gf_annotate(
    "label",
    label = "Rating = mu_Friends + Δ_ModernFamily + Δ_BigBangTheory + e",
    y = max(ratings_long$Rating) - 0.5,
    x = 2.25,
    size = 3,
    color = "black",
    fill = "lightblue"
  ) %>%
  gf_annotate(
    "label",
    label = "Triangles are Predictions\nat each Show level",
    y = min(ratings_long$Rating) + 0.5,
    x = 1.25,
    size = 3,
    color = "black",
    fill = "moccasin"
  ) %>%
  gf_labs(
    title = "Fitted ANOVA Model Equation",
    x = "TV Show",
    y = "Rating"
  ) 

```
### Inference

- The actual ratings for Big Bang Theory show the widest spread (more variation) compared to the other two shows.

## Normality Test

```{r}
shapiro.test(x = ratings_long$Rating)
```

```{r}
ratings_long %>%
  group_by(Show) %>%
  group_modify(~ .x %>%
    select(Rating) %>%
    as_vector() %>%
    shapiro.test() %>%
    broom::tidy())
```

```{r}
ratings_anova$residuals %>%
  as_tibble() %>%
  gf_dhistogram(~value, data = .) %>%
  gf_labs(
    title = "Residuals Histogram",
    x = "Residuals", y = "Count"
  ) %>%
  gf_fitdistr()

ratings_anova$residuals %>%
  as_tibble() %>%
  gf_qq(~value, data = .) %>%
  gf_qqstep() %>%
  gf_labs(
    title = "Residuals Q-Q Plot",
    x = "Theoretical Quantiles", y = "Sample Quantiles"
  ) %>%
  gf_qqline()

```

```{r}
ratings_long %>%
  group_by(Gender) %>%
  summarise(variance = var(Rating))

ratings_long %>%
  group_by(Show, Gender) %>%
  summarise(variance = var(Rating))

```

```{r}
DescTools::LeveneTest(Rating ~ Show, data = ratings_long)
```
### Inference

- The p-value (0.33) is greater than 0.05, so there’s no big difference in variances.

- The spread of ratings is similar across all three shows.

```{r}
fligner.test(Rating ~ Show, data = ratings_long)
```

```{r}
ratings_supernova$Show
```

```{r}
ggstatsplot::ggbetweenstats(
  data = ratings_long,
  x = Show,
  y = Rating,
  colour = Show,
  alpha = 0.8,
  type = "parametric",
  p.adjust.method = "bonferroni",
  conf.level = 0.95,
  violin.args = list(width = 0.0)  # Removes violin plots like in frogs example
) +
  scale_colour_paletteer_d("ggthemes::colorblind") +
  labs(
    title = "ANOVA: Ratings by TV Show",
    x = "TV Show",
    y = "Rating",
    subtitle = "",
    caption = ""
  )
```
### Inference

- Highest Rating: Modern Family has the highest average score (8.30).

- Lowest Rating: Big Bang Theory has the lowest average score (6.45).

- Most Variation: Friends has the widest scatter of individual ratings, showing the most diverse opinions).

- Clear Favorite: Modern Family is significantly preferred over both Friends and Big Bang Theory.

- Sample Size: All three shows have the same number of ratings (46).

## Conclusion

- Modern Family has the highest average rating among the three shows (8.30).

- Friends follows, and Big Bang Theory has the lowest ratings (6.45).

- The pairwise tests indicate that Modern Family is significantly different (and higher) than both Friends and Big Bang Theory.

- Gender does not significantly influence ratings. The difference between males and females is not statistically significant.

- The residual plots show that the data is fairly normal. This means the ANOVA results are valid and reliable. 

- Levene’s and Fligner’s tests confirm that the variances are similar across the shows, meeting ANOVA’s assumptions. 

- Overall, viewers clearly prefer Modern Family, and the test shows significant differences in how people rate the three shows.

--------------------------------------------------------------------------------
