---
title: "A3 Tattoo Dataset"
author: "Swetha KV"
date: "2025-10-31"
categories: [A3]
format: html
---
# Do tattoos make people look attractive or not?

## Setting up R Packages

```{r}
library(tidyverse)
library(mosaic)
library(vcd)
library(visStatistics)
library(resampledata)
library(ggformula)
library(infer)
library(openintro)
```

## Read Data

```{r}
tattoo <- read.csv("tattoo.csv")
glimpse(tattoo)
```

## View Missing Data

```{r}
visdat::vis_dat(tattoo)
```

## Summerise: Examine Data

```{r}
tattoo %>% count(Tattoo_Attractive)
```

```{r}
tattoo %>% count(Gender)
```

## Data Munging

```{r}
tattoo_modified <- tattoo %>%
  dplyr::select(Gender, Tattoo_Attractive) %>%
  tidyr::drop_na(c(Gender, Tattoo_Attractive )) %>%
  mutate(
    Gender = factor(Gender,
                    levels = c("F", "M"),
                    labels = c("Female", "Male")
    ),
    Tattoo_Attractive = factor(Tattoo_Attractive,
                              levels = c("Yes", "No"),
                              labels = c("Yes", "No")
    )
  )

glimpse(tattoo_modified)

```

## Chi Sqaure Test

```{r}
contingency_table <- vcd::structable(Gender ~ Tattoo_Attractive,
  data = tattoo_modified
)
contingency_table
```

```{r}
gss_vcd_table <- vcd::structable(Gender ~ Tattoo_Attractive, data = tattoo_modified)
gss_vcd_table %>%
  vcd::mosaic(
    gp = shading_hsv, direction = "v",
    main = "Tattoo Attraction by Gender Mosaic Chart",
    legend = TRUE,
    labeling = labeling_border(
      varnames = c("F", "F"),  # Adjusts variable name labels
      rot_labels = c(90, 0, 0, 0),  # Rotates top-side labels
      just_labels = c(
        "left",   # Top side
        "left",   # Right side
        "left",   # Bottom side
        "right"   # Left side
      )
    )
  )
```
### Inference

- The chart and test results strongly suggest there is no association between gender and tattoo attraction in your dataset.

- Statistically, the proportions of Yes/No responses for tattoo attraction are about the same for both males and females.

```{r}
# Somewhat strange notation with visstat
chisq_visstat <- visstat(
  x = tattoo_modified, # dataframe
  y = "Gender", # y variable
  "Tattoo_Attractive"
) # x variables

chisq_visstat %>% summary()
```

### Hypothesis

Null Hypothesis: There is no association between gender and finding tattoos attractive.

Alternative Hypothesis: Null Hypothesis is true 

```{r}
chisq_visstat$observed
```

```{r}
chisq_visstat$expected
```

```{r}
chisq_visstat$stdres
```
Need more samples!

```{r}
summary(tattoo_modified$Gender)
summary(tattoo_modified$Tattoo_Attractive)
```

```{r}
contingency_table
```
```{r}
contingency_table <- table(tattoo_modified$Gender, tattoo_modified$Tattoo_Attractive)
```

```{r}
xq_test_object <- xchisq.test(contingency_table)
```

```{r}
xq_test_object %>%
  broom::tidy() %>%
  select(statistic) %>%
  as.numeric() -> X_squared_observed
X_squared_observed
```

```{r}
df <- 1
X_squared_observed <- 0
X_squared_critical <- qchisq(0.95, df)

mosaic::xqchisq(
  p = 0.95, df = df,
  return = c("plot"), verbose = FALSE,
  system = "gg"
) %>%
  gf_labs(
    x = "F value", title = "Critical and Observed Chi-Square Values"
  ) %>%
  gf_vline(
    xintercept = X_squared_observed,
    color = "red", linewidth = 1
  ) %>%
  gf_vline(
    xintercept = X_squared_critical,
    color = "dodgerblue",
    linewidth = 1
  ) %>%
  gf_annotate(
    x = X_squared_observed - 2.5, y = 0.15,
    geom = "label", label = "Observed\nChi-Square",
    fill = "red", alpha = 0.3
  ) %>%
  gf_annotate("curve", X_squared_observed - 2.5,
    y = 0.135, yend = 0.10,
    xend = X_squared_observed - 0.5,
    linewidth = 0.5, curvature = 0.3,
    arrow = arrow(length = unit(0.25, "cm"))
  ) %>%
  gf_annotate(
    # ***MODIFIED: Changed x from + 2.5 to + 1.5***
    x = X_squared_critical + 1.5, y = 0.15,
    geom = "label", label = "Critical\nChi-Square",
    fill = "lightblue"
  ) %>%
  gf_annotate("curve", X_squared_critical + 1.5,
    y = 0.135, yend = 0.10,
    xend = X_squared_critical + 0.25,
    linewidth = 0.5, curvature = -0.3,
    arrow = arrow(length = unit(0.25, "cm"))
  ) %>%
  gf_refine(
    scale_y_continuous(expand = expansion(add = c(0, 0.01))),
    scale_x_continuous(expand = expansion(add = c(3, 2)))
  )
 
```
### Inference

- This shows the expected frequency of different Chi-Square values under the null hypothesis (there is no significant relationship).




```{r}
tattoo_vcd_table <- vcd::structable(Tattoo_Attractive ~ Gender,
  data = tattoo_modified
)

vcd::mosaic(tattoo_vcd_table,
  gp = shading_max, legend = TRUE, direction = "v",
  labeling = labeling_border(
    varnames = c("F", "F"),
    rot_labels = c(90, 0, 0, 0),
    just_labels = c(
      "left",
      "left",
      "left",
      "right"
    )
  )
)

vcd::mosaic(tattoo_vcd_table,
  type = "expected", gp = shading_max, legend = TRUE, direction = "v",
  labeling = labeling_border(
    varnames = c("F", "F"),
    rot_labels = c(90, 0, 0, 0),
    just_labels = c(
      "left",
      "left",
      "left",
      "right"
    )
  )
)

vcd::assoc(tattoo_vcd_table,
  gp = shading_max, legend = TRUE, direction = "v",
  labeling = labeling_border(
    varnames = c("F", "F"),
    rot_labels = c(90, 0, 0, 0),
    just_labels = c(
      "left",
      "left",
      "left",
      "right"
    )
  )
)

```

### Inference

- The Chi-Square test was conducted to check the relationship between gender and whether people find tattoos attractive.

- The observed frequencies (Female: 15 Yes, 5 No; Male: 15 Yes, 5 No) are same to the expected frequencies.

- The Chi-Square value = 0, with 1 degree of freedom and a p-value = 1.0.

- Since the p-value (1.0) is greater than 0.05. This means there is no association between gender and perception of tattoo attractiveness.

- Both males and females responded in exactly the same way, showing no variation in opinion.

## Proportion Test

```{r}
tattoo_modified %>%
  group_by(Tattoo_Attractive) %>%
  count()
```

```{r}
tattoo_modified_clean <- tattoo_modified %>%
  drop_na(Tattoo_Attractive)
```

```{r}
tattoo_modified_clean %>%
  count(Tattoo_Attractive)
```

```{r}
tattoo_modified_clean %>%
  summarize(prop = mean(Tattoo_Attractive == "Yes"), n = n())
```

### Inference
- The test result confirms if this percentage is significantly above 50%.

```{r}

tattoo_modified_clean %>%
  drop_na() %>%
  gf_bar(~Tattoo_Attractive, fill = ~Gender) %>%
  gf_labs(
    x = "Found tattoos attractive?",
    title = "Proportion of Students Finding Tattoos Attractive"
  )

```
### Inference

- Majority of them find tattoos attractive and it's equally spread out

- Again no association between gender and perception of tattoo attractiveness

```{r}
mosaic::binom.test(~Tattoo_Attractive, data = tattoo_modified_clean, success = "Yes")

```

```{r}
mosaic::binom.test(~Tattoo_Attractive, data = tattoo_modified_clean, success = "Yes") %>%
  broom::tidy()
```

```{r}
tattoo_modified_clean %>%
  drop_na() %>%
  specify(response = Tattoo_Attractive, success = "Yes") %>%
  generate(reps = 999, type = "bootstrap") %>%
  calculate(stat = "prop") %>%
  get_ci(level = 0.95)
```

### Inference

- A sample proportion test was conducted to determine if the proportion of students who find tattoos attractive differs significantly from 0.5 (i.e., 50%).

- Out of 40 students, 30 (75%) said they find tattoos attractive.

- The sample proportion (p) = 0.75.

- The p-value obtained from the test is 0.0022, which is less than 0.05. This means that the proportion of students who find tattoos attractive is significantly different from 50%.

- The 95% confidence interval (0.588 to 0.873) indicates that the true proportion of students who find tattoos attractive likely falls between 59% and 87%.

## Conclusion

- The Chi sqaure result is Not Statistically Significant. We Fail to Reject the Null Hypothesis.

- However, because the sample size is small (only 40 participants), the test may not capture real-world differences. A larger sample is needed to draw stronger and more reliable conclusions.

- For now, we can only conclude that within this limited sample, gender has no impact on tattoo attractiveness.

- But 