---
title: "A3 Weekly Expense Dataset"
author: "Swetha KV"
date: "2025-11-01"
categories: [A3]
format: html
---

# Guys have more pocket money than Girls.

## Setting up R Packages

```{r}
library(tidyverse)
library(mosaic)
library(broom)
library(gt) 
library(visStatistics)
library(dplyr)
library(ggplot2)
library(infer)

```

## Read Data

```{r}
weekly_expense <- read.csv("weeklyexpence.csv")
glimpse(weekly_expense)
```

```{r}

weekly_expense <- weekly_expense %>%
  mutate(
    Name = str_squish(Name),
    Gender = str_squish(Gender)
  )

gender_summary <- weekly_expense %>%
  group_by(Gender) %>%
  summarise(
    Average_Expenditure = mean(Total_Expenditure_Last_Week, na.rm = TRUE),
    Total_Expenditure = sum(Total_Expenditure_Last_Week, na.rm = TRUE),
    Count = n()
  )

gender_summary

```

### Hypothesis 

- Null Hypothesis: There is no difference between the true mean weekly expenditure for Male and Female.

- Alternative Hypothesis: There is a significant difference between the mean weekly expenditure for Male and Female.

## Visualising data

```{r}
weekly_expense %>%
  gf_dhistogram(
    ~Total_Expenditure_Last_Week,
    fill = ~Gender,
    alpha = 0.5,
    title = "Weekly Expenditure by Gender"
  ) %>%
  gf_facet_grid(~Gender) %>%
  gf_fitdistr(dist = "dnorm") %>%
  gf_refine(scale_fill_brewer(palette = "Set2")) %>%
  gf_theme(theme(legend.position = "none"))
```
### Inference

- Most girls spend a small amount of money, and their spending is very consistent.

- The spending habits of boys are much more spread out and less consistent than the spending of girls.

- The highest individual spenders in the dataset are all boys.

```{r}
weekly_expense %>%
  gf_boxplot(
    Total_Expenditure_Last_Week ~ Gender, orientation = "x" ,
    fill = ~Gender
  ) %>%
  gf_refine(scale_fill_brewer(palette = "Set2"))
```

### Inference

- A few boys who spent a lot of money (outliers) made the average spending for all boys look much higher than it really is for most of them.

```{r}
weekly_expense %>%
  gf_jitter(
    Total_Expenditure_Last_Week ~ Gender,
    color = ~Gender,
    alpha = 0.6,
    title = "Individual Weekly Expenditures"
  )
```

```{r}
weekly_expense %>%
  group_by(Gender) %>%
  summarise(Average_Spend = mean(Total_Expenditure_Last_Week, na.rm = TRUE)) %>%
  gf_col(
    Gender ~ Average_Spend,
    fill = ~Gender,
    title = "Average Weekly Expenditure by Gender"
  ) %>%
  gf_refine(scale_fill_brewer(palette = "Set2"))

```

```{r}
weekly_expense %>%
  gf_violin(
    Total_Expenditure_Last_Week ~ Gender,
    fill = ~Gender,
    alpha = 0.7,
    title = "Distribution of Weekly Expenditure by Gender"
  ) %>%
  gf_refine(scale_fill_brewer(palette = "Set2"))

```

```{r}
obs_diff_expense <- diffmean(
  Total_Expenditure_Last_Week ~ Gender,
  data = weekly_expense,
  only.2 = FALSE
)
obs_diff_expense

```
- This is difference is huge because the male expense is not evenly distributed and because of few outliers the mean difference is high.

```{r}
shapiro.test(weekly_expense$Total_Expenditure_Last_Week[weekly_expense$Gender == "Female"])
shapiro.test(weekly_expense$Total_Expenditure_Last_Week[weekly_expense$Gender == "Male"])
```
### Inference

- The weekly expenditure data for both males and females is not normally distributed. This is also strongly supported by histogram chart.

```{r}
var.test(
  Total_Expenditure_Last_Week ~ Gender,
  data = weekly_expense
)
```
### Inference 

- There is a statistically significant difference in the variability (spread) of the weekly expenditures between males and females. 

- This also we saw in the graphs that Male spending is far more spread out and inconsistent than Female spending on the histogram chart.

```{r}
mosaic::t_test(
  Total_Expenditure_Last_Week ~ Gender,
  data = weekly_expense
) %>%
  broom::tidy() %>%
  gt() %>%
  tab_style(
    style = list(
      cell_fill(color = "orange"),
      cell_text(weight = "bold")
    ),
    locations = cells_body(columns = p.value)
  ) %>%
  tab_header(title = "t-test: Difference in Mean Weekly Expenditure by Gender")

wilcox.test(
  Total_Expenditure_Last_Week ~ Gender,
  data = weekly_expense,
  conf.int = TRUE,
  conf.level = 0.95
) %>%
  broom::tidy() %>%
  gt() %>%
  tab_style(
    style = list(
      cell_fill(color = "yellow"),
      cell_text(weight = "bold")
    ),
    locations = cells_body(columns = p.value)
  ) %>%
  tab_header(title = "Wilcoxon Test: Median Expenditure Difference by Gender")
```
### Inference

From t-Test
- 

From Mann-Whitney Test
- P-value 0.066 it's Borderline Significant

- Estimate is 1100 The median for boys is about 1,100 Rs. higher than the median for girls.


```{r}
obs_diff_expense_less <- diffmean(
  Total_Expenditure_Last_Week ~ Gender,
  data = weekly_expense
)

obs_diff_expense_less
```

```{r}
mosaic::t_test(Total_Expenditure_Last_Week ~ Gender,
  var.equal = FALSE,
  data = weekly_expense
) %>%
  broom::tidy()
```

```{r}
obs_diff_infer <- weekly_expense %>% 
  infer::specify(Total_Expenditure_Last_Week ~ Gender) %>%
  infer::calculate(
    stat = "diff in means",
    order = c("Male", "Female")
  )
obs_diff_infer
```


```{r}
null_dist <- weekly_expense %>%
  specify(Total_Expenditure_Last_Week ~ Gender) %>%
  hypothesize(null = "independence") %>%
  generate(reps = 5000, type = "permute") %>% 
  calculate(
    stat = "diff in means",
    order = c("Male", "Female") 
  )
null_dist

null_dist %>%
  visualise() + # Note this plus sign!
  shade_p_value(obs_diff_infer,
    direction = "two-sided"
  ) +
  scale_y_continuous(expand = c(0, 0))
```


```{r}
null_dist %>%
  get_p_value(
    obs_stat = obs_diff_expense_less,
    direction = "two_sided"
  )
```

```{r}
null_dist %>%
  infer::get_confidence_interval(
    point_estimate = obs_diff_expense_less,
    level = 0.95
  )
```

### Inference 
- The p-value (0.056) is slightly above the level 0.05. T the observed difference in average weekly expenditure between males and females is not statistically significant.


## Conclusion

- Boys Spend More! Overall, boys spend a higher amount of money than girls do.

- Boys Spending is Messy (t-Test): Boys' spending habits are very unpredictable (some spend very little, a few spend huge amounts), but girls' spending is much more consistent.

- Outliers Confuse Tests: A few boys who spent giant amounts (the outliers) made the standard average tests (like the t-test) unreliable and unable to clearly prove the difference.

- Mann-whitney test that ignores those extreme high amounts, it confirms that the difference in spending between the genders is real with difference 1100 Rs.

--------------------------------------------------------------------------------

