---
title: "A2 Map Dataset"
author: "Swetha, Shruthi"
date: "2025-10-18"
categories: [Crane Mangolia]
format: html
---

This dataset explores the migration patterns of Cranes in and across Mangolia. The dataset contains various attributes includes the longitude and latitude positions of the cranes documented along their migration which we can use to better understand and track their migration patterns at certain times of the year.

```{r}
library(rnaturalearth)
library(rnaturalearthdata)

# Run this in your console first
# devtools::install_github("ropensci/rnaturalearthhires")
library(rnaturalearthhires)

# Plotting Maps
library(tidyverse) # Maps using ggplot() + geom_sf()
library(ggplot2)
library(ggformula) # Maps using gf_sf()
library(tmap) # Thematic Maps, static and interactive
library(tmaptools)
library(tmap.mapgl)
library(osmdata) # Fetch map data from osmdata.org
library(sfheaders) # Handcrafted Map data
## Interactive Maps
library(leaflet) # interactive Maps
library(leaflet)
library(leaflet.providers)
library(leaflet.extras)
library(threejs) # Globe maps in R. Part of the htmlwidgets family of packages

# For Spatial Data Frame Processing
library(sf)
```

```{r}
cranes <- readr::read_csv("cranesmangolia.csv") %>% 
janitor::clean_names(case="snake")
cranes
```

```{r}
glimpse(cranes)
```


```{r}
crane_location <- cranes %>%
  select(location_long,location_lat)

glimpse(crane_location)
```

```{r}
visdat::vis_miss(crane_location)
```

```{r}
# https://boundingbox.klokantech.com/ CSV:
bbox_1 <- matrix(c(65.3,19.2,122.0,53.8), byrow = FALSE, nrow = 2,
    ncol = 2, dimnames = list(c("x", "y"), c("min", "max")))
bbox_1
```

```{r}
cranes_sf <- st_as_sf(cranes,
                      coords = c("location_long", "location_lat"),  # check column names
                      crs = 4326)

```

```{r}
st_geometry(cranes_sf)
```
```{r}
craneslines <- st_read("lines.shp")
```

```{r}
cranespoints <- st_read("points.shp")
```

```{r}

migration_countries_1 <- World %>%
  filter(iso_a3 %in% c(
    "AFG", "CHN", "IND", "KAZ", "KGZ", 
    "MNG", "NPL", "PAK", "RUS", "TUR", 
    "TKM", "UZB"
  ))

# Plot with gf
gf_sf(migration_countries_1, fill = "grey", color = "black") %>%
  gf_point(data = cranes, location_lat ~ location_long, color = "violet", size = 1) %>%
  gf_labs(
    title = "Demoiselle Crane Migration Countries and Locations",
    x = "Longitude",
    y = "Latitude"
  )


```

```{r}

migration_countries <- World %>%
  filter(iso_a3 %in% c(
    "AFG", "CHN", "IND", "KAZ", "KGZ", 
    "MNG", "NPL", "PAK", "TUR", 
    "TKM", "UZB"
  ))

# Plot with gf
gf_sf(migration_countries, fill = "grey", color = "black") %>%
  gf_point(data = cranes, location_lat ~ location_long, color = "violet", size = 1) %>%
  gf_labs(
    title = "Zoom in -Demoiselle Crane Migration Countries and Locations",
    x = "Longitude",
    y = "Latitude"
  )


```

```{r}
gf_sf(migration_countries, fill = "grey", color = "black") %>%
gf_sf(data = craneslines, geometry = ~geometry, color = "darkblue", linewidth = 0.05) %>%
  gf_sf(data = cranespoints, geometry = ~geometry, color = "lightblue", alpha = 0.2, size = 0.5) %>%
  gf_theme(axis.text.x = element_text(angle = 45, hjust = 2)) %>%
  gf_labs(
    title = "CDemoiselle Crane Migration Countries and Locations ",
    subtitle = "Cranes Mangolia Study Source: Movebank"
  )
```
- Inference
1. Demoiselle Cranes move from northern Central Asia to South Asia. Most fixes cluster around Mongolia, China, and India. One outlier is far south.

2. The biggest crowd of cranes is in India. Probably they go there during winter season.

```{r}
gf_sf(migration_countries, fill = "grey", color = "black") %>%
  gf_sf(data = craneslines, geometry = ~geometry, color = "white", linewidth = 0.05) %>%
  gf_point(data = cranes, location_lat ~ location_long, color = ~ground_speed, size = 0.5) %>%
  gf_labs(
    title = "Demoiselle Crane Migration with Ground Speed",
    x = "Longitude",
    y = "Latitude",
    color = "Ground Speed (m/s)"
  )

```
-Inference
1. The color of the dots shows their ground speed. Darker colors represents slower speed, and lighter shades represents faster speed.

2. Most points are in darker shades. Thus, the cranes usually fly at slower speeds during migration.

```{r}
gf_sf(migration_countries, fill = "grey", color = "black") %>%
  gf_sf(data = craneslines, geometry = ~geometry, color = "white", linewidth = 0.05) %>%
gf_point(data = cranes, location_lat ~ location_long, color = ~height_above_ellipsoid) %>% 
    gf_labs(
    title = "Demoiselle Crane Migration with Height",
    x = "Longitude",
    y = "Latitude",
    color = "GHeight above ellipsoid"
  )
```
-Inference
1. They seems to fly high when they have to cross mountain otherwise, they fly flat and slow probably to save energy.

```{r}
gf_sf(migration_countries, fill = "grey", color = "black") %>%
  gf_point(data = cranes,location_lat ~ location_long, color = ~individual_local_identifier , size = 1)

```


```{r}
gf_sf(migration_countries, fill = "white", color = "black") %>%
  gf_line(data = cranes, location_lat ~ location_long, color = ~individual_local_identifier) %>%
    gf_point(data = cranes,location_lat ~ location_long, color = "black" , size = 0.2)
```

```{r}
print(craneslines)
plot(craneslines)
```

-Inferences
1. Most cranes travel together.

2. Each color is a different bird.

3. Some birds take slightly different paths.

4. A few birds stop at different places.
