---
title: "A3 Weird Dad Dataset"
author: "Swetha KV"
date: "2025-11-01"
categories: [Weird Dad]
format: html
---

Are Srishti dads weird?

## Setting up R Packages

```{r}
library(tidyverse)
library(mosaic)
library(vcd)
library(visStatistics)
library(resampledata)
library(ggformula)
library(infer)
library(openintro)
```

## Read Data

```{r}
weirddad <- read.csv("wierddad.csv")
glimpse(weirddad)
```

## Data Dictionary

## Cleaning Data

```{r}
weirddad <- weirddad%>% 
 tidyr::drop_na()
```

```{r}
visdat::vis_dat(weirddad)
```

## Data Munging

```{r}
weirddad_clean <- weirddad %>%
  dplyr::select(college, is_dad_weird) %>%
  tidyr::drop_na(c(college, is_dad_weird)) %>%
  mutate(
    college = factor(college),
    is_dad_weird = factor(is_dad_weird, levels = c("Yes", "No"))
  )
```

```{r}
glimpse(weirddad_clean)
```

## Summaries: Examining the Data

```{r}
weirddad_clean %>% count(is_dad_weird)
```

```{r}
weirddad_clean %>% count(college)

```

# Visualing data

```{r}
weirddad_clean %>%
  drop_na() %>%
  gf_bar(~college, fill = ~is_dad_weird, position = "fill") %>%
  gf_labs(
    x = "Is your dad weird?",
    title = "Propotion of SMI vs MIT saying their dad is weird"
  )
```

### Inference

-   more SMI students think their dad is weird than MIT students.

```{r}
weirddad %>%
  drop_na() %>%
  gf_bar(~is_dad_weird, fill = ~gender, position = "fill") %>%
  gf_labs(
    x = "Is your dad weird?",
    title = "Proportion of Students saying their dad is weird - Gender wise"
  )
```

### Inference

-   Gender does not appear to influence whether students think their dad is weird.

-   Males and females show almost equal proportions saying “Yes” or “No.”

-   This pattern shows the idea that perception of dad weirdness is gender-neutral — both male and female students perceive their fathers similarly.

## Chi Sqaure Test

```{r}
contingency_table <- vcd::structable(college ~ is_dad_weird,
  data = weirddad_clean
)
contingency_table

```

At MIT, 5 out of 33 students (15%) said their dad is weird. At SMI, 12 out of 31 students (39%) said their dad is weird.

```{r}
gss_vcd_table <- vcd::structable(is_dad_weird ~ college, data = weirddad_clean)
gss_vcd_table %>%
  vcd::mosaic(
    gp = shading_hsv, direction = "v",
    main = "College vs 'Is Dad Weird?' Mosaic Chart",
    legend = TRUE
      )
```

```{r}
contingency_table <- vcd::structable(college ~ is_dad_weird, data = weirddad_clean)
contingency_table <- as.table(contingency_table)
xq_test_object <- xchisq.test(contingency_table)
```

```{r}
xq_test_object %>%
  broom::tidy() %>%
  select(statistic) %>%
  as.numeric() -> X_squared_observed
X_squared_observed
```

```{r}
# Determine the Chi-Square critical value
X_squared_critical <- qchisq(
  p = .05,
  df = (5 - 1) * (2 - 1), # (nrows-1) * (ncols-1)
  lower.tail = FALSE
)
X_squared_critical
```

```{r}
test_result <- chisq.test(contingency_table)

# Extract required values
X_squared_observed <- test_result$statistic
df <- test_result$parameter
X_squared_critical <- qchisq(0.95, df)
```

```{r}
mosaic::xqchisq(
  p = 0.95, df = df,
  return = c("plot"), verbose = F,
  system = "gg"
) %>%
  gf_labs(
    x = "Chi-Square Value", title = "Critical and Observed Chi-Square Values",
    x = ""
  ) %>%
  gf_vline(
    xintercept = X_squared_observed,
    color = "red", linewidth = 1
  ) %>%
  gf_vline(
    xintercept = X_squared_critical,
    color = "dodgerblue", linewidth = 1
  ) %>%
  gf_annotate(
    x = X_squared_observed - 1, y = 0.15,
    geom = "label", label = "Observed\nChi-Square",
    fill = "red", alpha = 0.3
  ) %>%
  gf_annotate("curve", X_squared_observed - 1,
    y = 0.135, yend = 0.10,
    xend = X_squared_observed - 0.3,
    linewidth = 0.5, curvature = 0.3,
    arrow = arrow(length = unit(0.25, "cm"))
  ) %>%
  gf_annotate(
    x = X_squared_critical + 1, y = 0.15,
    geom = "label", label = "Critical\nChi-Square",
    fill = "lightblue"
  ) %>%
  gf_annotate("curve", X_squared_critical + 1,
    y = 0.135, yend = 0.10,
    xend = X_squared_critical + 0.3,
    linewidth = 0.5, curvature = -0.3,
    arrow = arrow(length = unit(0.25, "cm"))
  ) %>%
  gf_refine(
    scale_y_continuous(expand = expansion(add = c(0, 0.01))),
    scale_x_continuous(expand = expansion(add = c(0, 1)))
  )
```

## Propotion Test

```{r}
weirddad_clean %>%
  group_by(is_dad_weird) %>%
  count()
```

```{r}
glimpse(weirddad_clean)
```

```{r}
weirddad_clean %>%
  summarize(prop = mean(is_dad_weird == "Yes"), n = n())
```

### Inference

-   The difference in the proportion of students saying “Yes, my dad is weird” in college is statistically significant.

-   Across both colleges (MIT and SMI combined), about 26.6% of students said their dad is weird.

-   This means that most students (≈73%) do not consider their dad weird, suggesting that “weird dads” are a minority perception overall.

-   MIT: 15% “weird dads” SMI: 39% “weird dads”

-   That contrast shows that SMI is well above the overall average (26.6%), while MIT is below it.

-   SMI dads are perceived as significantly weirder than MIT dads.

```{r}
weirddad_clean %>%
  drop_na() %>%
  gf_bar(~is_dad_weird,  position = "dodge") %>%
  gf_labs(
    x = "Is your dad weird?",
    title = "Proportion of Students saying their dad is weird"
  )
```

```{r}
mosaic::binom.test(~is_dad_weird, data = weirddad_clean, success = "Yes")

```

```{r}
mosaic::binom.test(~is_dad_weird, data = weirddad_clean, success = "Yes") %>%
  broom::tidy()
```

```{r}
weirddad_clean %>%
  drop_na() %>%
  specify(response = is_dad_weird, success = "Yes") %>%
  generate(reps = 999, type = "bootstrap") %>%
  calculate(stat = "prop") %>%
  get_ci(level = 0.95)
```

------------------------------------------------------------------------
